\subsection{Learning $\nfhef$}

We now describe a learning algorithm for $\nfhef$.
Our algorithm is based on Angluin's $L^*$ algorithm \cite{Angluin87} for regular languages, and aims to learn a minimal deterministic permutation-sequence complete $\nfhef$ with a minimal number of variables for a hyperlanguage $\cal L$.

As with $L^*$, our algorithm consists of two entities: a {\em learner} and a {\em teacher}.
During the learning process, the learner asks the teacher two types of queries: {\em membership queries} (``is the hyperword $S$ in $\cal L$?'') and {\em equivalence queries} (``Is $\A$ an $\nfhef$ for $\cal L$?'').


The number of variables, as well as the quantification condition $\alpha$, is unknown to the learner in advance, and is established during the learning procedure. 
In every step, the learner may cause one of the following to occur in the $\nfhef$ $\A$ that it maintains:
\begin{enumerate}
    \item Increase the number of states;
    \item Increase the number of variables;
    \item Turn an $\exists$ quantifier to a $\forall$ quantifier.
\end{enumerate}

As with $L^*$, the learner maintains an {\em observation table} $T$, which, intuitively, maintains the states and transitions of the underlying automaton that have been learned so far. In addition, the learner maintains the current number of variables $k$ (which also influences the alphabet of the observation table), and the number of $\exists$ quantifiers $t$ in the quantification condition. 

In case of a failed equivalence query, the teacher returns a {\em counterexample}: a {\em negative counterexample} is a hyperword that is not in $\lang{\A}$ and is in $\cal L$, and a {\em positive counterexample} is a hyperword that is in $\lang{\A}$, but is not in $\cal L$. We assume that the teacher always returns a minimal counterexample (in terms of set size).

The actions of the learner depend on the properties of the returned counterexample $S$ it gets from the teacher. 
We say that $S$ is {\em manifested} in $T$ if there exists a word $w$ in $T$ that is formed by zipping the words in $S$ in some order.
\begin{enumerate}
    \item 
    If $S$ is a positive counterexample of size $k$ or less, that is not manifested in $T$, then the learner adds  $\zip(S)$ to $T$, where $\zip(S)$ denotes zipping the words in $S$ in some arbitrary order. 
    \item 
    If $S$ is a positive counterexample of size $k$ or less, that is manifested in $T$, then the learner sets $t\leftarrow t-1$. 
    \item
    If $S$ is a positive counterexample of size $k'>k$, then the learner replaces every letter $\sigma$ in the transitions with $\uparrow^{k'}(\sigma)$, and sets $t\leftarrow t+(k'-k)$. 
    \item If $S$ is a negative counterexample, and $k>t$. then it must be that $S$ is of size $k$ or less, and does not already appear in $T$. The learner then adds $S$ to $T$.
    \item if $S$ is a negative counterexample of size $k'>k$, and $k=t$ (that is, the quantifiers are all $\exists$), then the learner sets $t\leftarrow t-1$.  
\end{enumerate}




\input{figures/learning_flow}

